Parsing Tweets into Universal Dependencies

### 2017-12-05 paper.tex
updates and comments.

\begin{document}
\maketitle
\begin{abstract}
We study the problem of analyzing tweets with
universal dependencies \citep[UD;][]{NIVRE16.348}. We extend the UD guidelines to cover
special constructions in tweets. Using the extended guidelines \yicomment{including tokenization, POS tagging and parsing? (need to mention the annotation process/pipeline system not only consists parsing in abstract)}, we create
a new tweet treebank ({\sc Tweebank v2}), which has 3,550 tweets\yicomment{add tweet \#} and 55,607 tokens with
labeled attachments. 
It is more than four times larger than the (unlabeled) {\sc Tweebank
  v1} introduced by \citet{kong-EtAl:2014:EMNLP2014}. 
We characterize the disagreement
among our annotations and show that it is challenging to deliver
consistent annotation due to ambiguity in
understanding and explaining tweets. Using the new treebank,
we build a pipeline system to parse raw tweets into UD. To overcome the
annotation noise without sacrificing computational efficiency, we propose a new
method to distill an ensemble of 20 transition-based parsers into a single one. Our
parser achieves an improvement of 2.6 in LAS over \nascomment{which
  comparison is this?}\yjcomment{my bad, it was fixed.} the baseline
and outperforms parsers that are state-of-the-art on other treebanks in both accuracy and speed.
\end{abstract}

@@

The annotation effort we describe stands in contrast to \yicomment{the?} previous work
by \newcite{kong-EtAl:2014:EMNLP2014}.  Their aim was the rapid
development of a dependency parser for tweets, and to that end they
contributed a new annotated corpus, \textsc{Tweebank}, consisting of
929 tweets.  Their annotations added \yicomment{unlabelled?} dependencies to a portion of the data
annotated with POS tags by 
\newcite{gimpel-EtAl:2011:ACL-HLT2011} and
\newcite{owoputi-EtAl:2013:NAACL-HLT} after rule-based tokenization \citep{ICWSM101540}.
Kong et al.~also contributed a system for parsing;
we defer the discussion of their parser to \S\ref{sec:parsing}.

@@

In the different direction of splitting tokens, UD guidelines also suggest to merge
{\it multi-token words} (e.g. {\it 20 000}) into one single token in some special
cases. We witnessed a small number of tweets that contain multi-token words
(e.g. {\it Y O}, and {\it R E T W E E T}) but didn't combine them due to simplification.
Such tokens only account for 0.067\% and we use the UD
  {\it goeswith} relation to resolve these cases in the dependency annotations.

@@

Tokens that do not have a syntactic function (discussed at greater
length in the next section) were usually annotated as \emph{other}
(\texttt{X}), except for emoticons, which are tagged as \emph{symbol}
(\texttt{SYM}), following UD\_English, seen in Figure \ref{fig:non-syn-toks} as an example.

@@

Another important difference follows from the difference in
tokenization.  As discussed in \S\ref{sec:tok-anno}, UD calls for more
aggressive tokenization than that of \citet{ICWSM101540} which opted
out of tokenizing contractions and possessives and it was used by 
\citet{gimpel-EtAl:2011:ACL-HLT2011}. As a consequence, 
Gimpel et al. introduced new parts of speech for these cases
instead.\footnote{These tags only account for 2.7\% of tokens,
  leading to concerns about data sparseness in tagging and all
  downstream analyses.}  For us, these tokens must be split, but
universal parts of speech can be applied.

@@

\paragraph{Acronym abbreviations.}  How should we syntactically
analyze acronym tokens like \emph{idc}
(abbreviating ``I don't care'')
 and \emph{rn} (``right now'')?  Should
they be decomposed into their component words, and if so should those
words be ``normalized'' into explicitly spelled out intermediate
forms? \yicomment{this sentence should be deleted or moved to tokenization?} \yicomment{As we already know from previous sections, \emph{idc} is not tokenized and tagged as verb, we follow ...} \yicomment{Normalization is discussed in the 2.5.2 Annotation Process. Actually we did do normalization, but just as auxiliary info, didn't use it for any model.}
We follow \citet{kong-EtAl:2014:EMNLP2014} and annotate their
syntax as a single word without normalization. Their syntactic functions
are decided according to their context. \citet{eisenstein:2013:NAACL-HLT} studied the necessity of normalization
in social media text and argued that such normalization is problematic.
Our solution to the syntax of abbreviations follows the spirit of his argument. Because abbreviations which clearly carry
syntactic functions only constitute 0.06\% of the tokens in our
dataset, we believe that normalization is an unnecessarily complicated step.\yicomment{Again we did normalization!}
% \nascomment{maybe we owe citations to
%   \cite{finin-EtAl:2010:MTURK,eisenstein:2013:NAACL-HLT} here?  maybe
%   better papers to cite}

@@

Inspired by the way UD deals with \emph{punctuation} (which is
canonically non-syntactic), we adopt the following
conventions:\yicomment{shall we mention how we segment sentences? Seems relevant here}

\begin{itemize}
%\item For the at-mentions in retweet constructions, we attach them to their retweet markers. We will discuss our choice in the following paragraph;
\item If one syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate. Retweet construction is a special case and we will discuss our choice in the following paragraph
\item If the whole sentence is made of a sequence of non-syntactic tokens, we attach all these tokens to the first one;
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English dataset.
%\\ \yicomment{For retweet at-mention case, we attach the retweet at-mentions to the RT sign, so that the retweet action and the people whose tweet is being retweeted could be connected.\footnote{We can also follow the first convention to attach the retweet at-mention to its predicate, but we think it is better for downstream tasks.}}
\end{itemize}

@@

\paragraph{Retweet construction.} Figure \ref{fig:non-syn-toks} shows
an example of the retweet construction (\emph{RT @coldplay :}).  This
might be treated as a verb phrase, with \emph{RT} as a verb and the
at-mention as its object\yicomment{or might be treated as 2 non-syntactic tokens and a punctuation, which should all attach to the predicate of the following retweeted tweet.}.  This solution would lead to an uninformative root
word and, since this expression is idiomatic to Twitter, might create
unnecessary confusion for downstream applications aiming to identify
the main predicate(s) of a tweet.  We therefore treat the whole
expression as non-syntactic, including assigning the \emph{other}
(\texttt{X}) part of speech to both \emph{RT} and \emph{@coldplay},
attaching the at-mention to \emph{RT} with the \emph{discourse}
label and the colon to \emph{RT} with the \emph{punct}(uation) label,
and attaching \emph{RT} to the predicate of the following sentence.
%\yicomment{since we have introduced retweet at-mentions in the non-syntactic tokens, here we can just briefly justify our choice.}.

@@

\paragraph{Vocative at-mentions.}  Another idiomatic construction on
Twitter is an at-mention %(followed by a colon \nascomment{?})
as a sign that a tweet is a reply to a tweet by the mentioned user\yicomment{not necessary, can be just mentioning somebody. should be careful to say it. We can say this construction include 1. reply;  2. mention sb. Both cases are similar to general vocative proper noun case in standard texts. And we need to exclude at-mentions in Retweet Construction.}.  We
treat these at-mentions as vocative expressions, labeling them as
\emph{proper noun} (\texttt{PROPN})\yicomment{here is confusing that PROPN could be thought as a label in dependencies.} and attaching them to the
following predicate with the label \emph{vocative} (see Figure \ref{fig:informal-toks} for an example) \yicomment{just as UD guidelines}.

@@

\subsubsection{Annotation Process}
Our annotation process was conducted in three stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the \yicomment{general} UD
annotation conventions and our guidelines \yicomment{specifically for annotating tweets}.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a tokenizer, a POS tagger, and
parser were trained on the annotated data from the first stage (1,050 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data. Finally, an extra
layer of word-level normalization\yicomment{see 10 th col of conllu data, also mention we could recover the original tweets also from this col (tokenization)} was manually annotated for data
analysis purposes. For example, {\em idc} is annotated as
`I\_don't\_care' and {\em 4 my FAM} is as `for my familiy'.
Such normalization doesn't change our tokenization annotation. \yicomment{should mention tokenization process, say we tokenize V1 and train tokenizer for V2 data, then manually correct them.}\yjcomment{add ``a tokenizer''}
\yicomment{add final data split} We obtained 3,550 labelled annotated tweets, split into train/dev/test set with /(639 $+$ 1000)/(210 $+$ 500)/(201 $+$ 1000) tweets, where each set consists of both old tweets from \citet{gimpel-EtAl:2011:ACL-HLT2011} and new tweets from the Twitter stream. \yicomment{should also mention main annotators (authors) are non-native English speakers?}

@@

Results of the \citet{owoputi-EtAl:2013:NAACL-HLT}  with non-greedy
inference on automatically tokenized data
are shown in  Table \ref{tbl:pos-result-vs-tok}.  We see that errors
in tokenization do propagate, but tagging performance is above 93\%
with our tokenizer. \yicomment{our POS tagger results on automatically tokenized data?}

@@

\paragraph{Pipeline evaluation.} Finally, we report  our full
pipeline's performance in  Table \ref{tbl:pipline}. We also compare
our model with a stack of the state-of-the-art systems (namely, the {\it SOTA stack}):
Stanford CoreNLP tokenizer\yicomment{why not compare with UD pipe?},
\citet{owoputi-EtAl:2013:NAACL-HLT}'s tagger, and \citet{DBLP:journals/corr/DozatM16}'s parser.
Our system differs with the SOTA stack in the tokenization and parser components.
From Table \ref{tbl:pipline}, our system outperforms the SOTA stack when
evaluated in pipeline manner.
The results also emphasize the
importance of segmentation\yicomment{you mean word tokenization or sentence segmentation?}:  without gold segmentation, UD parsing
performance drops by more than four points.

>>>>

\begin{document}
\maketitle
\begin{abstract}
We study the problem of analyzing tweets with
universal dependencies \citep[UD;][]{NIVRE16.348}. We extend the UD guidelines to cover
special constructions in tweets. Using the extended guidelines \yicomment{including tokenization, POS tagging and parsing? (need to mention the annotation process/pipeline system not only consists parsing in abstract)}, we create
a new tweet treebank ({\sc Tweebank v2}), which has 3,550 tweets\yicomment{add tweet \#}\yjcomment{didn't see necessity of mentioning \# of tweets} and 55,607 tokens with
labeled attachments. 
It is more than four times larger than the (unlabeled) {\sc Tweebank
  v1} introduced by \citet{kong-EtAl:2014:EMNLP2014}. 
We characterize the disagreement
among our annotations and show that it is challenging to deliver
consistent annotation due to ambiguity in
understanding and explaining tweets. Using the new treebank,
we build a pipeline system to parse raw tweets into UD. To overcome the
annotation noise without sacrificing computational efficiency, we propose a new
method to distill an ensemble of 20 transition-based parsers into a single one. Our
parser achieves an improvement of 2.6 in LAS over \nascomment{which
  comparison is this?}\yjcomment{my bad, it was fixed.} the baseline
and outperforms parsers that are state-of-the-art on other treebanks in both accuracy and speed.
\end{abstract}

@@

The annotation effort we describe stands in contrast to \yicomment{the?} previous work
by \newcite{kong-EtAl:2014:EMNLP2014}.  Their aim was the rapid
development of a dependency parser for tweets, and to that end they
contributed a new annotated corpus, \textsc{Tweebank}, consisting of
12,149 tokens.  Their annotations added \yicomment{unlabelled?} dependencies to a portion of the data
annotated with POS tags by 
\newcite{gimpel-EtAl:2011:ACL-HLT2011} and
\newcite{owoputi-EtAl:2013:NAACL-HLT} after rule-based tokenization \citep{ICWSM101540}.
Kong et al.~also contributed a system for parsing;
we defer the discussion of their parser to \S\ref{sec:parsing}.

@@

In the different direction of splitting tokens, UD guidelines also suggest to merge
{\it multi-token words} (e.g. {\it 20 000}) into one single token in some special
cases. We witnessed a small number of tweets that contain multi-token words
(e.g. {\it Y O}, and {\it R E T W E E T}) but didn't combine them for simplification.
Such tokens only account for 0.067\% and we use the UD
  {\it goeswith} relation to resolve these cases in the dependency annotations.

@@

Tokens that do not have a syntactic function (see Figure \ref{fig:non-syn-toks}, discussed at greater
length in the next section) were usually annotated as \emph{other}
(\texttt{X}), except for emoticons, which are tagged as \emph{symbol}
(\texttt{SYM}), following UD\_English.

@@

Another important difference follows from the difference in
tokenization.  As discussed in \S\ref{sec:tok-anno}, UD calls for more
aggressive tokenization than that of \citet{ICWSM101540} which opted
out of splitting contractions and possessives. As a consequence of adopting \citet{ICWSM101540}'s tokenization, 
Gimpel et al. introduced new parts of speech for these cases
instead.\footnote{These tags only account for 2.7\% of tokens,
  leading to concerns about data sparseness in tagging and all
  downstream analyses.}  For us, these tokens must be split, but
universal parts of speech can be applied.

@@

\paragraph{Acronym abbreviations.}  How should we syntactically
analyze acronym tokens like \emph{idc}
(abbreviating ``I don't care'')
 and \emph{rn} (``right now'')?  Should
they be decomposed into their component words, and if so should those
words be ``normalized'' into explicitly spelled out intermediate
forms? \yicomment{this sentence should be deleted or moved to tokenization?} \yicomment{As we already know from previous sections, \emph{idc} is not tokenized and tagged as verb, we follow ...} \yicomment{Normalization is discussed in the 2.5.2 Annotation Process. Actually we did do normalization, but just as auxiliary info, didn't use it for any model.}\yjcomment{i think acronym is a different story to tokenization.}
We follow \citet{kong-EtAl:2014:EMNLP2014} and annotate their
syntax as a single word without normalization. Their syntactic functions
are decided according to their context. \citet{eisenstein:2013:NAACL-HLT} studied the necessity of normalization
in social media text and argued that such normalization is problematic.
Our solution to the syntax of abbreviations follows the spirit of his argument. Because abbreviations which clearly carry
syntactic functions only constitute 0.06\% of the tokens in our
dataset, we believe that normalization is an unnecessarily complicated step.\yicomment{Again we did normalization!}
% \nascomment{maybe we owe citations to
%   \cite{finin-EtAl:2010:MTURK,eisenstein:2013:NAACL-HLT} here?  maybe
%   better papers to cite}

@@

Inspired by the way UD deals with \emph{punctuation} (which is
canonically non-syntactic), we adopt the following
conventions:\yicomment{shall we mention how we segment sentences? Seems relevant here}\yjcomment{i don't want to touch sentence segmentation, it will makes the paper more messy.}

\begin{itemize}
%\item For the at-mentions in retweet constructions, we attach them to their retweet markers. We will discuss our choice in the following paragraph;
\item If one syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate. Retweet construction is a special case and we will discuss our choice in the following paragraph;
\item If the whole sentence is made of a sequence of non-syntactic tokens, we attach all these tokens to the first one;
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English dataset.
%\\ \yicomment{For retweet at-mention case, we attach the retweet at-mentions to the RT sign, so that the retweet action and the people whose tweet is being retweeted could be connected.\footnote{We can also follow the first convention to attach the retweet at-mention to its predicate, but we think it is better for downstream tasks.}}
\end{itemize}

@@

\paragraph{Retweet construction.} Figure \ref{fig:non-syn-toks} shows
an example of the retweet construction (\emph{RT @coldplay :}).  This
might be treated as a verb phrase, with \emph{RT} as a verb and the
at-mention as its object\yicomment{or might be treated as 2 non-syntactic tokens and a punctuation, which should all attach to the predicate of the following retweeted tweet.}\yjcomment{i wouldn't like the readers to think in this direction.}.  This solution would lead to an uninformative root
word and, since this expression is idiomatic to Twitter, might create
unnecessary confusion for downstream applications aiming to identify
the main predicate(s) of a tweet.  We therefore treat the whole
expression as non-syntactic, including assigning the \emph{other}
(\texttt{X}) part of speech to both \emph{RT} and \emph{@coldplay},
attaching the at-mention to \emph{RT} with the \emph{discourse}
label and the colon to \emph{RT} with the \emph{punct}(uation) label,
and attaching \emph{RT} to the predicate of the following sentence.
%\yicomment{since we have introduced retweet at-mentions in the non-syntactic tokens, here we can just briefly justify our choice.}.

@@

\paragraph{Vocative at-mentions.}  Another idiomatic construction on
Twitter is an at-mention %(followed by a colon \nascomment{?})
as a sign that a tweet is a reply to a tweet by the mentioned user\yicomment{not necessary, can be just mentioning somebody. should be careful to say it. We can say this construction include 1. reply;  2. mention sb. Both cases are similar to general vocative proper noun case in standard texts. And we need to exclude at-mentions in Retweet Construction.}\yjcomment{i didn't see any difference between reply and mention, they all starts with a @username.}.  We
treat these at-mentions as vocative expressions, labeling them as
\emph{proper noun} (\texttt{PROPN})\yicomment{here is confusing that PROPN could be thought as a label in dependencies.} and attaching them to the
following predicate with the label \emph{vocative} (see Figure \ref{fig:informal-toks} for an example) \yicomment{just as UD guidelines}.

@@

\subsubsection{Annotation Process}
\begin{table}
	\centering
	\begin{tabular}{rrrr}
		\hline
		& train & dev & test \\
		\hline
		tweets 		&	1,639 		&	709		&	1,201\\
		& +1,000 & +709 & +1,000\\
		tokens 		&	24,753		&	11,742	&	19,112\\
		\hline
	\end{tabular}
\caption{Statistics of {\sc Tweebank v2}.
The second column shows the number of new tweets compared to
{\sc Tweebank v1}.\label{tbl:data-stat}
}
\end{table}
Our annotation process was conducted in three stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the \yicomment{general} UD
annotation conventions and our guidelines \yicomment{specifically for annotating tweets}.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a tokenizer, a POS tagger, and
parser were trained on the annotated data from the first stage (1,050 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data. Finally, an extra
layer of word-level normalization\yicomment{see 10 th col of conllu data, also mention we could recover the original tweets also from this col (tokenization)}\yjcomment{didn't get your point. if mentioning the normalization here introduce too much confusions, we would better remove these.} was manually annotated for data
analysis purposes. For example, {\em idc} is annotated as
`I\_don't\_care' and {\em 4 my FAM} is as `for my familiy'.
Such normalization doesn't change our tokenization annotation. \yicomment{should mention tokenization process, say we tokenize V1 and train tokenizer for V2 data, then manually correct them.}\yjcomment{add ``a tokenizer''}
\yicomment{add final data split}\yjcomment{it's better to use table \ref{tbl:data-stat}. what's more, i don't think we benefit from re-stating our data source.} We obtained 3,550 labelled annotated tweets, split into train/dev/test set with /(639 $+$ 1000)/(210 $+$ 500)/(201 $+$ 1000) tweets, where each set consists of both old tweets from \citet{gimpel-EtAl:2011:ACL-HLT2011} and new tweets from the Twitter stream. \yicomment{should also mention main annotators (authors) are non-native English speakers?}

@@

Results of the \citet{owoputi-EtAl:2013:NAACL-HLT}  with non-greedy
inference on automatically tokenized data
are shown in  Table \ref{tbl:pos-result-vs-tok}.  We see that errors
in tokenization do propagate, but tagging performance is above 93\%
with our tokenizer.

@@

\paragraph{Pipeline evaluation.} Finally, we report  our full
pipeline's performance in  Table \ref{tbl:pipline}. We also compare
our model with a stack of the state-of-the-art systems (namely, the {\it SOTA stack}):
Stanford CoreNLP tokenizer\yicomment{why not compare with UD pipe?}\yjcomment{i was thinking stanford tokenizer is more suitable, but i will get the udpipe result},
\citet{owoputi-EtAl:2013:NAACL-HLT}'s tagger, and \citet{DBLP:journals/corr/DozatM16}'s parser.
Our system differs with the SOTA stack in the tokenization and parser components.
From Table \ref{tbl:pipline}, our system outperforms the SOTA stack when
evaluated in pipeline manner.
The results also emphasize the
importance of segmentation\yicomment{you mean word tokenization or sentence segmentation?}\yjcomment{we didn't touch sentence seg.}:  without gold segmentation, UD parsing
performance drops by more than four points.
